---
title: "ID and dependencies"
author: "Marsel Scheer"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ID and dependencies}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
futile.logger::flog.threshold(futile.logger::WARN)

library(projectPlan)
suppressPackageStartupMessages(library(dplyr))
library(DT)
```

# Introduction

The columns *id* is used to identify tasks in the work plan. This allows to specify dependencies between
tasks. There can be logic dependencies where a task cannot start if another has not been finished, for instance
first comes the fundament, then the walls and then the roof. Other tasks may need to be executed
sequentially simply because because there are not enough resources.
For illustration purpose, we will create very small project plans, that focus on the corresponding aspect 
under considerations.

# *id* not unique 

```{r, echo=FALSE}
pA <- dplyr::tibble(
  project = c("A"), 
  section = c("0_prep", "1_design"), task = paste0("task", 1:2), id = paste0("i", 1:2), start = "TODAY", depends_on = "", deadline = NA,
  end = 1:2,
  progress = 0, 
  resource = "R1")
```

The intrinsic property of an *id* is that it is unique. Internally, the *id*-entry is combined with the project name.
Therefore if the same *id* is used in different projects it is unique.
However, especially at the beginning of a project when the plan is growing it could be handy to have one id for more than one tasks. 
Using the same *id* within the same *section* is possible. If this happens, then the corresponding rows are combined
together into one entry, i.e.:

* all entries in columns *depends_on, resource, task* are concatenated
* mean of the progress column is calculated
* minimum of the deadlines is calculated
* sum of the estimated days necessary is calculated
* if start and end date are given, then minimum of start and maximum of the end date are used for the combined entry
* if any task waiting, then the combined entry also gets a waiting status

Note that using the same *id* in different *sections* results in a warning

```{r, echo=FALSE}
pA$id <- "i1"
DT::datatable(pA)
```

```{r, fig.height=5, fig.width=10}
pA %>% 
  projectPlan::wrangle_raw_plan() %>% 
  projectPlan::calculate_time_lines() %>% 
  projectPlan::gantt_by_sections()
```

