---
title: "Project and Section column"
author: "Marsel Scheer"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Project and Section}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
futile.logger::flog.threshold(futile.logger::WARN)

library(projectPlan)
suppressPackageStartupMessages(library(dplyr))
library(DT)
```

# Introduction

The columns *project* and *section* can be used to group task together also in the gantt-chart. Furthermore,
the *project* column is necessary if one wants to specify dependencies between tasks of different projects.
For illustration purpose, we will create very small project plans, that focus on the corresponding aspect 
under considerations.

# *project* column

Usually it is not necessary to create the *project* column explicitly. Importing the a project plan with
*projectPlan::import_xlsx* will use the name of the sheet as the project name. However, if a *project*
column exists the name of the sheet is ignored.

Lets look at two very simple projects containing two tasks each:
```{r, echo=FALSE}
pA <- dplyr::tibble(
  project = c("A"), 
  section = "prep", task = paste0("task", 1:2), id = paste0("i", 1:2), start = "TODAY", depends_on = "", deadline = NA,
  end = 1:2,
  progress = 0, 
  resource = "R1")

pB <- pA
pB$project = "B"

pp <- dplyr::bind_rows(pA, pB)

DT::datatable(pp)
```

Creating a gantt-chart one sees that every plan gets its own blue frame:
```{r, fig.height=5, fig.width=10}
pp %>% 
  projectPlan::wrangle_raw_plan() %>% 
  projectPlan::calculate_time_lines() %>% 
  projectPlan::gantt_by_sections()
```

Usually, dependencies between tasks can be specified by using the entry in the *id* column. However, to reference
an id from another project you have to use *::*, similar to the notation that is necessary if you want to explicitly
state the namespace in *R*. In order to specify that id *i1* of project *A* depends on *i1* of project *B* one enters
*B::i1*. Note, that we need to remove the *start*-time for the first entry, otherwise the *start*-time would be used
and not the time calculated due to the dependency.

```{r, echo=FALSE, fig.height=5, fig.width=10}
pp$depends_on[1] = "B::i1"
pp$start[1] = NA
DT::datatable(pp)
pp %>% 
  projectPlan::wrangle_raw_plan() %>% 
  projectPlan::calculate_time_lines() %>% 
  projectPlan::gantt_by_sections(show_dependencies = TRUE)
```
This example also shows that the same *id*'s can be used in different projects.

# *section* column

Within every blue frame, i.e. a project, of a gantt-chart further dotted frames/lines separate the different sections within that project,
where the section are sorted alphabetically.

```{r, echo=FALSE}
pA <- dplyr::tibble(
  project = c("A"), 
  section = c("0_prep", "1_design"), task = paste0("task", 1:2), id = paste0("i", 1:2), start = "TODAY", depends_on = "", deadline = NA,
  end = 1:2,
  progress = 0, 
  resource = "R1")

DT::datatable(pA)
```

```{r, fig.height=5, fig.width=10}
pA %>% 
  projectPlan::wrangle_raw_plan() %>% 
  projectPlan::calculate_time_lines() %>% 
  projectPlan::gantt_by_sections()
```


