---
title: "Start and end dates"
author: "Marsel Scheer"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Start and end dates}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
logger::log_threshold(logger::WARN)

library(projectPlan)
suppressPackageStartupMessages(library(dplyr))
library(DT)
```

# Introduction

The duration of a task is defined by the two columns *start* and *end*. The duration of the task
can be implicitly (due to dependencies and estimated durations of a task) or explicitly defined.

# Explicit start and end dates

This is the easiest case. Just provide the date in the format *yyyy-mm-dd* for start and end date.
If you use a Excel dates are stored as numeric numbers. In order convert this numeric back to a 
date a origin is necessary, which can be passed to *projectPlan::wrangle_raw_plan*. 
Unfortunately, Excel uses two different origins. For dates (post-1901) from Windows Excel origin 
should by 1899-12-30 (is the default value). For dates (post-1901) from Mac Excel origin should 
by 1904-01-01. 

It is recommended to set the start date to the actual date when the task was started. 
This way it is directly documented when the task was started. It is also recommend to set the
actual end date once the task was completed.

```{r}
pA <- dplyr::tibble(
  project = c("A"), 
  section = "prep", task = paste0("task", 1:2), id = paste0("i", 1:2), 
  start = c("2018-12-10", "2018-12-11"),
  end   = c("2018-12-15", "2018-12-12"),
  progress = 0, 
  resource = "R1")
DT::datatable(pA)
```

# Implicit start and end dates

Lets look at a simple dependency, where id *i2* depends on id *i1*, which it self is implicitly defined by the *TODAY*-entry 
in column *start*:
```{r}
pA <- dplyr::tibble(
  project = c("A"), 
  section = "prep", task = paste0("task", 1:2), id = paste0("i", 1:2), 
  start = c("TODAY", NA_character_),
  end   = c(NA_character_, NA_character_),
  depends_on = c(NA_character_, "i1"),
  est_duration = c(2,4),
  deadline = c(NA_character_, NA_character_),
  status = c("", ""),
  progress = 0, 
  resource = "R1")
DT::datatable(pA)
```

Wrangling this simple plan allows to calculate the timelines for both tasks:

```{r}
pA %>% 
  projectPlan::wrangle_raw_plan() %>% 
  calculate_time_lines() %>% 
  select(task, time_start, time_end) %>% 
  DT::datatable()
```

Note that *TODAY* for the start time of id *i1* was subsituted by the date the function *wrangle_raw_plan*
was executed.

See the corresponding vignette for more details on id's and dependencies.
