---
title: "Start and end dates"
author: "Marsel Scheer"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Start and end dates}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
logger::log_threshold(logger::WARN, namespace = "projectPlan")

library(projectPlan)
suppressPackageStartupMessages(library(dplyr))
library(DT)
```

# Introduction

The duration of a task is defined by the two columns *start* and *end*, which 
can be implicitly (due to dependencies and estimated durations of a task) or explicitly defined.

# Explicit start and end dates

This is the easiest case. Just provide the date in the format *yyyy-mm-dd* for start and end date.
If you use a Excel dates are stored as numeric numbers. In order convert this numeric back to a 
date a origin is necessary, which can be passed to *projectPlan::wrangle_raw_plan*. 
Unfortunately, Excel uses two different origins. For dates (post-1901) from Windows Excel origin 
should be 1899-12-30 (is the default value). For dates (post-1901) from Mac Excel origin should 
be 1904-01-01. 

One possible workflow is to set the start date to the actual date when the task was started. 
This way it is directly documented when the task was started and also to set the
actual end date once the task was completed. 

```{r, echo = FALSE}
pA <- dplyr::tibble(
  project = c("A"), 
  section = "prep", task = paste0("task", 1:2), id = paste0("i", 1:2), 
  start = c("2018-12-10", "2018-12-11"),
  end   = c("2018-12-15", "2018-12-12"),
  progress = 0, 
  resource = "R1",
  microtasks = "-",
  comments = "-")
DT::datatable(pA)
```
All dates are interpreted as 'begin of business day' (BOB) which is just a conventation made 
for this package. Of course, it would also be reasonable
to define the end date as 'end of busibness day'. However, there is already a third date and
other dates might follow. Hence, it is probably easier to always think in BOB. So a task that
started on 2018-12-10 and ended on 2018-12-11 was accomplished within one day.


# Implicit start and end dates

Lets look at a simple dependency, where id *i2* depends on id *i1*:
```{r, echo = FALSE}
pA <- dplyr::tibble(
  project = c("A"), 
  section = "prep", task = paste0("task", 1:2), id = paste0("i", 1:2), 
  start = NA_character_,
  end   = NA_character_,
  depends_on = c(NA_character_, "i1"),
  est_duration = c(2,4),
  deadline = NA_character_,
  status = c("", ""),
  progress = 0, 
  resource = "R1",
  microtasks = "-",
  comments = "-")
DT::datatable(pA)
```

Wrangling this simple plan allows to calculate the timelines for both tasks:

```{r}
pA %>% 
  projectPlan::wrangle_raw_plan() %>% 
  calculate_time_lines() %>% 
  select(task, time_start, time_end) %>% 
  DT::datatable()
```

Note that the start time of id *i1* is subsituted by today by the function 
*projectPlan::wrangle_raw_plan*.

See the corresponding vignette for more details on id's and dependencies.

# (Un)scheduled tasks

The package has a logic to determine if a task is unscheduled or not.
This is just a flag that helps to filter the tasks. The user can always
explicitly specify that a task as unscheduled (just write "unscheduled" 
into the status-column of the spreadsheet). However, it is a bit
cumbersome to maintain this status. Therefore the package automatically
flags a task as (un)scheduled if:

* no explicit start date is given and no prior task is defined, then the task is flagged as unscheduled 
* only if all ids given in the depends_on-column are flagged as scheduled, then the task is flagged as scheduled (irrespective of the status of the ids specified in the start-column)
* if any id in start-column is flagged as scheduled and the depends_on-column is empty, then the task is flagged as scheduled